---
title: "Good Software Engineering Practice for R Packages"
---

<!--
Creators (initial authors):
- Daniel Sabanes Bove, https://github.com/danielinteractive/ and www.linkedin.com/in/danielsabanesbove/
- Friedrich Pahlke, https://github.com/fpahlke/ and www.linkedin.com/in/pahlke/
- Kevin Kunzmann, https://github.com/kkmann/ and https://www.linkedin.com/in/kevin-kunzmann-6486a11bb/
- Liming Li, https://github.com/clarkliming
- Joe Zhu, https://github.com/shajoezhu and http://www.linkedin.com/in/joe-zhu-464b5818
- Shuang Li, https://github.com/shuangli22
- Doug Kelkhoff, https://github.com/dgkf and https://www.linkedin.com/in/doug-kelkhoff/
- Philippe Boileau, https://github.com/philboileau and https://www.linkedin.com/in/philippe-boileau-773270205/
- Lukas A. Widmer, https://github.com/luwidmer and https://www.linkedin.com/in/lukas-widmer-37800b9b/
- Michael Mayer, https://www.linkedin.com/in/drmichaelmayer/
- Louisa H. Smith, https://www.linkedin.com/in/louisahsmith/

The current version was edited by
- Daniel Sabanes Bove, https://github.com/danielinteractive/ and www.linkedin.com/in/danielsabanesbove/

License: In general unless otherwise stated, this work is licensed under the Creative Commons Attribution-ShareAlike 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by-sa/4.0/.
The source files are hosted at https://github.com/RCONIS/workshop-zurich-2026, which is forked from
the original version at https://github.com/openpharma/workshop-r-swe-mtl.
The Zurich 2026 revision also adapts updated chapters from
https://github.com/openstatsware/shortcourse-munich2026; author credits are preserved
in the individual chapters. Changes include course details, scheduling, roxygen2 tags,
SSH setup, and a new AI chapter.

Important: To use this work you must provide the name of the creators (initial authors), a link to the material, a link to the license, and indicate if changes were made.
-->

Welcome to the homepage of the workshop *"Good Software Engineering Practice for R Packages"*, part of the Zurich R Course series. In this course you will learn hands-on skills and tools to engineer reliable R packages used in statistics. The workshop will be conducted in two days and will be a mix of presentations and exercises. Participants should have at least one year of experience using R, be fluent in writing R functions, and bring their own laptop and charger.

## Event Details

The workshop takes place on **Thursday 8 and Friday 9 October 2026, 09:00–16:30**, in person at **ETH Zurich**, and is taught in **English**.
See the [official course information](https://math.ethz.ch/sfs/education/zurich-r-courses/r-packages.html) and [registration and venue page](https://sce.ethz.ch/en/programmes-and-courses/search-current-courses/wbk/software-eng-R-2026.html).

The instructor is **Daniel Sabanés Bové**, Director, inferential.biostatistics GmbH.
The course is part of [Zurich R Courses](https://math.ethz.ch/sfs/education/zurich-r-courses.html), the joint continuing education programme of UZH and ETH Zurich. Course management: Dr Lukas Meier, Seminar for Statistics, ETH Zurich.

The course fee is CHF 600 for members of UZH/ETH and associated institutes, and CHF 1,200 otherwise. Internet access is provided for software and package installation.

Topics: R package syntax, software engineering workflow, package quality, collaboration via GitHub, publication, Shiny development, and using AI assistants.

## Workshop Program

### Day 1: 8 October 2026

```{r}
#| echo: false
readr::read_csv("slides/resources/program.csv", col_types = "ccc") |>
  dplyr::filter(day == "1") |>
  dplyr::select(- day) |>
  knitr::kable(col.names = c("", ""))
```

### Day 2: 9 October 2026

```{r}
#| echo: false
readr::read_csv("slides/resources/program.csv", col_types = "ccc") |>
  dplyr::filter(day == "2") |>
  dplyr::select(- day) |>
  knitr::kable(col.names = c("", ""))
```

## Prerequisites & Technical Setup

Prior to the course, participants should

-   Set up a (free) [GitHub.com account](https://github.com/join).
    - Note: There are other git platforms like [Gitlab](https://about.gitlab.com/) or [Bitbucket](https://bitbucket.org/), but we made the choice to go with GitHub.com for the course, since it is by far the most relevant git platform in the R community.
-   Download and extract [simulatr.zip](slides/download/simulatr.zip)
-   Install the latest [R](https://www.r-project.org/) and [RStudio](https://posit.co/download/rstudio-desktop/) software.
-   Install [Rtools](https://cran.r-project.org/bin/windows/Rtools/) (on Windows; choose the version matching your R installation). On macOS, install the [build tools](https://mac.r-project.org/tools/) if needed to compile packages from source.
-   Install additional R packages using the [installation script](slides/download/install.R).

For the course, participants are required to use their own laptop to be able to participate in the exercises.

### Optional reading list

-   Excellent and very comprehensive [R Packages (2e)](https://r-pkgs.org/)
-   (Hardcore) description of how [Writing R Extensions](https://cran.r-project.org/doc/manuals/R-exts.html) works
-   [GitHub 'Hello World'](https://docs.github.com/en/get-started/quickstart/hello-world) tutorial on how to use GitHub.com (does not require git command line knowledge)
-   Tutorial on [version control with git and SVN](https://support.posit.co/hc/en-us/articles/200532077-Version-Control-with-Git-and-SVN) (we will be using git)
-   Command line git mini intro [trygit](https://trygit.js.org/)
-   [First steps with git & github by RStudio](https://www.youtube.com/watch?v=jN6tvgt3GK8)
-   [RStudio cheatsheets/ Git & GitHub](https://github.com/rstudio/cheatsheets/blob/main/git-github.pdf)
-   [Happy Git: SSH keys, including setup from RStudio](https://happygitwithr.com/ssh-keys.html#create-an-ssh-key-pair)
-   [Tidy design principles by Hadley Wickham](https://tidydesign.substack.com/)
-   [Posit skills for AI assistants](https://github.com/posit-dev/skills)

## Last events

| Date | Location | Materials |
| --- | --- | --- |
| 13 August 2026 | Munich, DE | [Workshop](https://openstatsware.github.io/shortcourse-munich2026/) |
| 17th July 2026 | Seoul, KR | [Workshop](https://openstatsware.github.io/shortcourse-seoul2026/) |
| 10th October 2025 | Paris, FR | [Workshop](https://rconis.github.io/SnB2025-shortcourse/) |
| 24th August 2025 | Basel, CH | [Workshop](https://openstatsware.github.io/shortcourse-iscb2025/) |
| 8th April 2025 | Tokyo, JP | [Workshop](https://www.linkedin.com/posts/yuichi-nakajima-8984759a_r-package-development-workshop-by-daniel-activity-7305102975851511810-QPl_?utm_source=share&utm_medium=member_desktop&rcm=ACoAAAeLiYUB2lCTxf2k2-ROB6XktxT7lyQYIaQ) |
| 8th July 2024 | Salzburg, AT | [Workshop](https://rconis.github.io/user2024-tutorial-gswep/) |
| 18th and 19th April 2024 | Zurich, CH | [Workshop](https://rconis.github.io/workshop-r-swe-zrh/) |
| 16th and 17th October 2023 | Montreal, QC, CA | [Workshop](https://openpharma.github.io/workshop-r-swe-mtl/) |
| 26th September 2023 | Rockville, MD, US | [Workshop](https://openpharma.github.io/workshop-r-swe-md/) |
| 20th July 2023 | San José, CA, US | [Workshop](https://openpharma.github.io/workshop-r-swe-sf/) |
| 24th March 2023 | Shanghai, CN | [Workshop](https://openpharma.github.io/workshop-r-swe/) |
| 10th February 2023 | Basel, CH | [Workshop](https://kkmann.github.io/workshop-r-swe/) |

## Credits and license

Unless otherwise credited, the course materials are licensed under [CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/). Initial and subsequent authors are listed in each slide chapter.
This Zurich 2026 revision by Daniel Sabanés Bové adapts the [Zurich 2024 course](https://github.com/RCONIS/workshop-r-swe-zrh) and the updated [Munich 2026 materials](https://github.com/openstatsware/shortcourse-munich2026), which followed the [Seoul 2026 course](https://github.com/openstatsware/shortcourse-seoul2026).
Changes include course information, scheduling, documentation tags, SSH setup, and the Using AI chapter. See [attribution notes](ATTRIBUTION.md) for the full source history and third-party image credit.
